---
title: "A kasstylesr example workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A kasstylesr example workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# Introduction to kasstylesr

## The following presents a tutorial to using kasstylesr to help automate your chart-building workflow

### Setup:
```{r setup}
library(kasstylesr)
library(ggplot2)
library(scales)
```

### Using color_picker() and kas_style() for formatting:
``` {r}
# Grab some example data
df <- iris
```

### The following uses color_picker(3) to set the colour palette, kas_style() is also added to provide formatting
``` {r, fig.width=8, fig.height=4.5}
chart_1 <- ggplot(df, aes(Petal.Width, Petal.Length, colour = Species)) +
  geom_point() +
  # kas_style() is also added to the plot to help with formatting
  kas_style() +
  # color_picker can take an integer as input
  scale_colour_manual("",
    values = color_picker(3),
    # set the legend order manually
    breaks = c("virginica", "versicolor", "setosa")
  ) +
  geom_text(x = 1.4, y = 6.2, label = "virginica", color = "black") +
  geom_text(x = 0.7, y = 4, label = "versicolor", color = "black") +
  geom_text(x = 0.8, y = 1.5, label = "setosa", color = "black") +
  theme(legend.position = "none") +
  xlab("Petal Width (cm)") +
  ylab("Petal Length (cm)")

# enact the code
chart_1
```

### While the following takes a list of strings (your categories) as input
``` {r, fig.width=8, fig.height=4.5}
types <- c("virginica", "versicolor", "setosa")
chart_2 <- ggplot(df, aes(Petal.Width, Petal.Length, colour = Species)) +
  geom_point() +
  kas_style() +
  scale_colour_manual("",
    values = color_picker(types),
    # set the legend order manually
    breaks = types
  ) +
  geom_text(x = 1.4, y = 6.2, label = "virginica", color = "black") +
  geom_text(x = 0.7, y = 4, label = "versicolor", color = "black") +
  geom_text(x = 0.8, y = 1.5, label = "setosa", color = "black") +
  theme(legend.position = "none") +
  xlab("Petal Width (cm)") +
  ylab("Petal Length (cm)")

# enact the code
chart_2
```


### Formatting bar charts works in the same way:
``` {r, fig.width=8, fig.height=4.5}
country <- c(
  rep("United Kingdom", 3), rep("Australia", 3),
  rep("United States", 3), rep("Sweden", 3)
)
occupation <- rep(c("Child", "Working Age", "OAP"), 4)
value <- abs(rnorm(12, 1000000000, 100000000))
data <- data.frame(country, occupation, value)

chart_3 <- ggplot(data, aes(fill = occupation, y = value, x = country)) +
  geom_bar(position = "fill", stat = "identity") +
  # the primary colour palette is used for bar charts
  scale_fill_manual(values = color_picker(3, primary_palette = TRUE)) +
  kas_style() +
  ylab("occupation share of each country") +
  scale_y_continuous(
    labels = scales::label_percent()
  ) +
  theme(axis.text.x = element_text(angle = 90))

# enact the code
chart_3
```

color_picker() will also provide live good-practice warnings. Attempting to write a stacked bar chart with a primary palette and 4 categories gives a warning and defaults to the multicolour palette
``` {r, fig.width=8, fig.height=4.5}
country <- c(
  rep("U.K.", 4), rep("Australia", 4), rep("U.S.", 4),
  rep("Sweden", 4)
)
occupation <- rep(c("Child", "Working Age", "Unemployed", "OAP"), 4)
value <- abs(rnorm(16, 1000000000, 100000000))
data <- data.frame(country, occupation, value)

chart_4 <- ggplot(data, aes(fill = occupation, y = value, x = country)) +
  geom_bar(position = "fill", stat = "identity") +
  # the primary colour palette only has 3 colours,
  # so trying to use 4 will prompt
  # a return to the multi-colour palette
  scale_fill_manual(values = color_picker(4, primary_palette = TRUE)) +
  kas_style() +
  ylab("occupation share of each country") +
  scale_y_continuous(
    labels = scales::label_percent()
  )

# enact the code
chart_4
```

### Saving charts for publication

Prepare the items required into lists below, for each ggplot chart, we need 1 title, source 
 and alt text to be declared. Every list must be in chart order. ie
 the 1st item in alt_texts_list will feed into the chart built in the 1st Excel tab,
which uses the chart declared in the first value of charts_list

``` {r}
# list of ggplot chart object to add to the release
charts_list <- list(chart_1, chart_2, chart_3, chart_4)
# list of names you want your charts to be saved as and/or
# the tab names for the excel workbook
chart_names_list <- c("Chart 1", "Chart 2", "Chart 3", "Chart 4")
# list of chart titles(contain each in strings('')),
# which appear above the chart in the excel workbook
# Ensure the chart titles use 'to' instead of a hyphen
# when referring to 2 time periods.
titles_list <- titles_list <- c(
  "Petal dimensions by species", "Petal dimensions by species",
  "Demographics by country", "Demographics by country"
)
# list of data souces you want displayed beneath the chart
# (contain each in strings(''))
sources_list <- c(
  "Source: R Iris dataset", "Source: R Iris dataset",
  "Source: Example data", "Source: Example data"
)
# list of alt texts(contain each in strings('')),
# which appear underneath the chart in the excel workbook
alt_texts_list <- c(
  "Alt text: This describes the chart's meaning", "Alt text",
  "Alt text", "Alt text"
)
```

### The following calls the functions to save and build the chart workbook

the following functions saves the list of plots to the folder "charts", which will be created automatically
``` {r}
project_area = './'
save_list_of_charts_to_folder(chart_names_list, charts_list, project_area)
```
The following generates the path to the 'charts' folder created above, no edit required

project_area is the same as specified above, the function will automatically find the 'charts' folder created
``` {r}
workbook_name <- 'vignette_charts.xlsx'
chart_folder_to_workbook(
  project_area, workbook_name, chart_names_list,
  titles_list, sources_list, alt_texts_list
)
```
